# Reproducible Research: Peer Assessment 1

```{r setoptions, echo = FALSE}
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

Loading the data:

```{r loaddata}
data <- read.csv(unz("activity.zip", "activity.csv"), sep = ",", header = TRUE)
data$date <- as.Date(data$date)
```

## What is mean total number of steps taken per day?

Missing values in the dataset are ignored.

```{r histogramsteps}
hist(data$steps, main = "Histogram of the total number of steps taken each day", xlab = "Number of steps")
```

Mean total number of steps taken per day:
```{r stepsmean}
steps.mean <- mean(data$steps, na.rm = TRUE)
steps.mean
```

Median total number of steps taken per day:
```{r stepsmedian}
steps.median <- median(data$steps, na.rm = TRUE)
steps.median
```

## What is the average daily activity pattern?

Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days:

``` {r plot}
x <- levels(as.factor(data$interval))
y <- sapply(x, function(i) {mean(data$steps[data$interval == i], na.rm = TRUE)})
plot(x, y, type = "l")
```

Calculating, which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps:
```{r maxsteps}
max.steps <- x[match(max(y), y)]
max.steps
```

## Imputing missing values

Total number of missing values in the dataset (i.e. the total number of rows with NAs):
```{r numberofNAs}
missing <- sum(!complete.cases(data))
missing
```

Missing values in the dataset are filled in with means for given 5-minute intervals.
```{r fill}
fdata <- data
for (i in seq(nrow(data))) {
    if (is.na(data$steps[i])) {
        fdata$steps[i] <- mean(data$steps[data$interval == data$interval[i]], na.rm = TRUE)
    }
}
```

Histogram of the total number of steps taken each day:
```{r fhistogramsteps}
hist(fdata$steps)
```

Mean total number of steps taken per day:
```{r fstepsmean}
steps.mean <- mean(fdata$steps)
steps.mean
```

Median total number of steps taken per day:
```{r fstepsmedian}
steps.median <- median(fdata$steps)
steps.median
```

## Are there differences in activity patterns between weekdays and weekends?

Creating a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day:
```{r weektime}
fdata$weektime <- as.factor(ifelse(as.POSIXlt(fdata$date)$wday %in% c(0,6),"weekend", "weekday"))
```

```{r fplot}
par(mfrow=c(2,1))
fx1 <- levels(as.factor(fdata$interval[fdata$weektime=="weekday"]))
fy1 <- sapply(fx1, function(i) {mean(fdata$steps[fdata$weektime=="weekday" & fdata$interval == i])})
fx2 <- levels(as.factor(fdata$interval[fdata$weektime=="weekend"]))
fy2 <- sapply(fx2, function(i) {mean(fdata$steps[fdata$weektime=="weekend" & fdata$interval == i])})
plot(fx1, fy1, type = "l", main = "Weekday", xlab = "Interval", ylab = "Number of steps")
plot(fx2, fy2, type = "l", main = "Weekend", xlab = "Interval", ylab = "Number of steps")
```
